---
title: "Analysis - Gender Equity"
author: "Chelsea Ursaner"
date: "February 2, 2017"
output:
  pdf_document: default
  html_document: default
---

**Background:** [EXECUTIVE DIRECTIVE ON GENDER EQUITY IN CITY OPERATIONS](http://www.lamayor.org/mayor-garcetti-issue-executive-directive-gender-equity-city-operations)

**Objective:** Analyze 2/2 employee database extract to identify areas of gender bias in City gov hiring based on Executive Directive areas of focus. 

Read data
```{r}
setwd("~/R_Datasets/DataLA")
data <- read.csv("employee_demographics.csv")
```

**Data Prepping**
```{r}
#See a sample of the data, column formats, and overall size of the dataset
library(dplyr)
glimpse(data)

#perform transformations to improve dataset 
#filter to only full time employees
data <- filter(data, LEGACY_DESCRIPTION == "Full-Time Regular")
#remove the dollar sign from hourly base rate so R reads it as numeric
library(stringr)
data$HOURLY_BASE_RATE <- as.numeric(str_replace_all(data$HOURLY_BASE_RATE, "\\$", ""))
library(lubridate) #fix date formats as needed
data$ORIGINAL_HIRE_DATE <- mdy(data$ORIGINAL_HIRE_DATE)
data$EMPLOYMENT_DATE <- mdy(data$EMPLOYMENT_DATE)
```

**Enrichment**
Create new groupings/variables in order to understand female representation in different jobs

```{r}
#Aggregating to look at female representation by job (3 options)
#get counts of females and males - looking at just representation
data$Female <- ifelse(data$GENDER == "Female",1,0)
data$Male <- ifelse(data$GENDER == "Male",1,0)
```

**Exploration and Visualization**
How do the findings change depending on how I define a job?

```{r}
#Option 1 - Job and department
jobs1 <- data %>% group_by(PAYROLL_DEPT_DESC, JOB_CLASS_TITLE) %>%
  summarize(CountF = sum(Female),
            CountM = sum(Male),
            Total = (sum(Female) + sum(Male)),
            PercentFemale = CountF/Total)
library(ggplot2)
sum(data$Female)/nrow(data)
p1 <- qplot(jobs1$PercentFemale,
            main = "Level of detail: Job and department. I.e. Mgmt assistant in LAWA",
            ylab = "Number of Records",
            xlab = "Percent Female")


#Option 2 - Job and department, filtering to jobs that have more than 8 people
jobs2 <- filter(jobs1, Total > 5)
p2 <- qplot(jobs2$PercentFemale,
            main = "Same as #1 but filtered to jobs that have >8 people",
            ylab = "Number of Records",
            xlab = "Percent Female")

#Option 3 - just job class title and filtering to > 10 people
jobs3 <- data %>% group_by(JOB_CLASS_TITLE) %>%
  summarize(CountF = sum(Female),
            CountM = sum(Male),
            Total = (sum(Female) + sum(Male)),
            PercentFemale = CountF/Total) %>%
  filter(Total > 10)

p3 <- qplot(jobs3$PercentFemale,
            main = "LOD Job Class (i.e. Mgmt Assistant) filtered to classes with >10 people",
            ylab = "Number of Records",
            xlab = "Percent Female")

library(gridExtra)
grid.arrange(p1, p2, p3)
```


**Results/Next Steps**
Creating two lists - one of jobs that are heavily male-dominated and another of jobs that are heavily female-dominated. From here, I'll work with the 

```{r}
#Using option 3, get subsets of gender biased jobs
male_bias <- jobs3 %>% filter(PercentFemale < .1)
female_bias <- jobs3 %>% filter(PercentFemale > .9)

#Display as tables
library(knitr)
kable(female_bias, digits = 2, caption = 'Female-Dominated Jobs (where women represent > 90%)')
kable(male_bias, digits = 2, caption = 'Male-Dominated Jobs (where men represent > 90%)')

#as new files
write.csv(male_bias, "male_bias.csv", row.names = F)
write.csv(female_bias, "female_bias.csv", row.names = F)
```




